---
title: "Untitled"
format: html
editor: visual
---

Simulate data for Linear quantile regression

Heteroskedastic error with increasing variance simulation

OLS

```{r eval = T, warning = F, message = F}
# Library section
library(quantreg)
library(ggplot2)
library(tidyverse)
library(sn)
```

```{r eval = T, warning = F, message = F}
set.seed(479)
sample_size <- c(50, 100, 200, 450, 700)
repli <- 1000
beta0_original <- 15
beta1_original <- 0.4
sigma_error <- 2.3
mean_error <- 0
beta_error <- 1/7
```

```{r eval = T, warning = F, message = F}
set.seed(479)
ols_qr_heteroscedastic <- list()
slope_func <- function(data, indices) {
  d <- data[indices, ]
  fit <- lm(y ~ x, data = d)
  coef(fit)[2]
}
```

```{r}
library(lmtest)
library(sandwich)
x <- runif(300, min = 0, max = 5)
    y <- beta0_original + beta1_original * x + rnorm(300, mean_error, sigma_error) * (1 + x * beta_error)
dat <- data.frame(x = x, y = y)

    ols <- lm(y ~ x, data = dat)
    
    qr <- rq(y ~ x, tau = tau, data = dat)
    summary(qr, se = "nid")
```

```{r}
# Looping areaaaaaaaa
for (i in seq_along(sample_size)) {
  n <- sample_size[i]
  
  coverage <- logical(repli)
  slope <- numeric(repli)
  se_slope <- numeric(repli)
  bias <- numeric(repli)
  power <- logical(repli)

  for (j in 1:repli) {
    # data set creation
    x <- runif(n, min = 0, max = 5)
    y <- beta0_original + beta1_original * x + rnorm(n, mean_error, sigma_error) * (1 + x * beta_error)
    dat <- data.frame(x = x, y = y)
    
    # OLS
    ols <- lm(y ~ x, data = dat)
    slope[j] <- coef(ols)[2]
    se_slope[j] <- coeftest(ols)$coefficients["x", "Std. Error"]
    
    # Build 95% confidence interval
    ci_lower <- slope[j] - 1.96 * se_slope[j]
    ci_upper <- slope[j] + 1.96 * se_slope[j]
    
    coverage[j] <- (ci_lower <= beta1_original) && (ci_upper >= beta1_original)
    power[j] <- (summary(ols)$coefficients["x", "Pr(>|t|)"] < 0.05)
    bias[j] <- slope[j] - beta1_original
  }
  results <- tibble(
    Run = 1:repli,
    Sample_Size = n,
    Model = "OLS",
    Slope = round(slope, 4),
    Bias = bias,
    SE_Slope = round(se_slope, 4),
    Coverage = coverage,
    Power = power,
    CI_Lower = round(slope - 1.96 * se_slope),
    CI_Upper = round(slope + 1.96 * se_slope),
    True_Slope = beta1_original
  )
  
  ols_qr_heteroscedastic[[i]] <- results
}
ols_qr_heteroscedastic <- bind_rows(ols_qr_heteroscedastic)

head(ols_qr_heteroscedastic, n = 3)
```

```{r eval = T, warning = F, message = F}
set.seed(479)
taus <- c(0.25, 0.5, 0.75, 0.9)
qr_heteroscedastic <- list()

for (tau in taus) {
  cat("Quantile =", tau, "\n")
  beta1.true <- beta1_original + qnorm(tau) * sigma_error * beta_error
  cat("True slope =", beta1.true, "\n")
 
  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
    coverage <- logical(repli)
    slope <- numeric(repli)
    se_slope <- numeric(repli)
    power <- logical(repli)
   
    for (j in 1:repli) {
      x <- runif(n, min = 0, max = 5)
      y <- beta0_original + beta1_original * x + rnorm(n, mean_error, sigma_error) *(1 + x * beta_error)
     
      dat <- data.frame(x = x, y = y)
      qr <- rq(y ~ x, tau = tau, data = dat)
     
      slope[j] <- coef(qr)[2]
     
      sum_rq <- summary(qr, se = "nid")
      se_slope[j] <- sum_rq$coefficients["x", "Std. Error"]
    
      ci_lower <- slope[j] - 1.96 * se_slope[j]
      ci_upper <- slope[j] + 1.96 * se_slope[j]
     
      coverage[j] <- (ci_lower <= beta1.true) && (ci_upper >= beta1.true)
      power[j] <- !(ci_lower <= 0 && ci_upper >= 0)
      bias[j] <- slope[j] - beta1.true
    }
    
    results <- tibble(
      Run = 1:repli,
      Sample_Size = n,
      Model = paste("QR", tau, sep = "_"),
      Slope = round(slope, 4),
      Bias = bias,
      SE_Slope = round(se_slope, 4),
      Coverage = coverage,
      Power = power,
      CI_Lower = round(slope - 1.96 * se_slope),
      CI_Upper = round(slope + 1.96 * se_slope),
      True_Slope = round(beta1.true, 4)
    )
    
    qr_heteroscedastic <- append(qr_heteroscedastic, list(results))
  }
}
ols_qr_heteroscedastic <- bind_rows(ols_qr_heteroscedastic, bind_rows(qr_heteroscedastic))

ols_qr_heteroscedastic[seq(1, nrow(ols_qr_heteroscedastic), by = 1000), ]
```

```{r}
models <- unique(ols_qr_heteroscedastic$Model)

for (mod in models) {
  
  df_mod <- filter(ols_qr_heteroscedastic, Model == mod)
  ci_lower_avg <- mean(df_mod$CI_Lower)
  ci_upper_avg <- mean(df_mod$CI_Upper)
  True_Slope <- mean(df_mod$True_Slope)
  Bias <- mean(df_mod)
  slope_all_heter <- ggplot(ols_qr_heteroscedastic, aes(x = factor(Sample_Size), y = Slope, fill = Model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(data = distinct(ols_qr_heteroscedastic, Model, True_Slope), 
             aes(yintercept = True_Slope, color = Model)) +
  labs(title = "Estimated vs True Slopes",
       x = "Sample Size", y = "Estimated Slope") +
  theme_minimal() +
  theme(legend.position = "bottom")

  saveRDS(slope_all_heter, "slope_all_heter.rds")

  if (mod == "OLS") {
    p <- ggplot(df_mod, aes(x = factor(Sample_Size), y = Slope)) +
    geom_boxplot(alpha = 0.7, outlier.alpha = 0.3, fill = "steelblue") +
    geom_hline(aes(yintercept = mean(Slope)), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
    annotate("rect", xmin = -Inf, xmax = Inf,
             ymin = ci_lower_avg,
             ymax = ci_upper_avg,
             fill = "blue",
             alpha = 0.07) +
    annotate("text", x = Inf, y = True_Slope + 0.5,
             label = paste("Average Slope =", round(True_Slope, 3)),
             hjust = 1, vjust = 0,color = "red", size = 4,
             fontface = "bold")
    labs(title = paste("Slope - ", mod), x = "Sample Size", y = "") +
    theme_minimal()
  } else {
    p <- ggplot(df_mod, aes(x = factor(Sample_Size), y = Slope)) +
    geom_boxplot(alpha = 0.7, outlier.alpha = 0.3, fill = "steelblue") +
    geom_hline(aes(yintercept = True_Slope), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
    annotate("rect", xmin = -Inf, xmax = Inf,
             ymin = ci_lower_avg,
             ymax = ci_upper_avg,
             fill = "blue",
             alpha = 0.07) +
    annotate("text", x = Inf, y = True_Slope + 0.5,
             label = paste("True Slope =", round(True_Slope, 3)),
             hjust = 1, vjust = 0,color = "red", size = 4,
             fontface = "bold") +
    labs(title = paste("Slope - ", mod),
         x = "Sample Size", y = "") +
    scale_fill_brewer(palette = "Set3") +
    theme_minimal()
  }
  print(p)
  saveRDS(p, paste0(mod,"_heters"))
}
```

```{r}
bias_heter <- ggplot(ols_qr_heteroscedastic, aes(x = factor(Sample_Size), y = Bias, fill = Model)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  labs(title = "Distribution of Slope Estimates by Sample Size", x = "Sample Size", y = "Bias")
print(bias_heter)
saveRDS(bias_heter, "Bias_all_heter")
```

```{r}
power_coverage_heter <- ols_qr_heteroscedastic %>%
  group_by(Sample_Size, Model) %>%
  summarise(Power_rate = mean(Power) * 100,
    Coverage_rate = mean(Coverage) * 100,
    .groups = "drop"
  )

power_heter <- ggplot(power_coverage_heter, aes(x = Sample_Size, y = Power_rate, color = Model)) +
  geom_line() +
  geom_hline(yintercept = 95, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("rect", 
           xmin = -Inf, xmax = Inf,
           ymin = 95 - 1.96 * sqrt(0.05 * 0.95/repli)*100,
           ymax = 95 + 1.96 * sqrt(0.05 * 0.95/repli)*100,
           alpha = 0.1, fill = "blue") +
  labs(title = "Distribution of Power Rate", x = "Sample Size", y = "Power percentage") +
  theme(legend.position = "none")

cover_heter <- ggplot(power_coverage_heter, aes(x = Sample_Size, y = Coverage_rate, color = Model)) +
  geom_line() +
  geom_hline(yintercept = 95, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("rect", 
           xmin = -Inf, xmax = Inf,
           ymin = 95 - 1.96 * sqrt(0.05 * 0.95/repli)*100,
           ymax = 95 + 1.96 * sqrt(0.05 * 0.95/repli)*100,
           alpha = 0.1, fill = "blue") +
  labs(title = "Distribution of Coverage Rate", x = "Sample Size", y = "Coverage percentage")

library(patchwork)
power_heter | cover_heter

saveRDS(power_heter, "power_heter")
saveRDS(cover_heter, "cover_heter")
```

## Skewed with constant variance simulation

OLS

```{r eval = T, warning = F, message = F}
# library(sn) # library of skew to normal dsitribution
library(sn)
set.seed(479)

skewness <- c(0, 2, 2.5, 2.7, 3, 6, 7)

ols_skew <- list()

for (skew in skewness) {
  cat("\nSkewness =", skew, "\n")
  
  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
    
    slope <- numeric(repli)
    bias <- numeric(repli)
    coverage <- logical(repli)
    interval_length <- numeric(repli)
    power <- logical(repli)
    absolute_percentage_bias <- numeric(repli)
    se_slope <- numeric(repli)
   
    for (j in 1:repli) {
      # Generate error
      # if (skew == 0) {
      #   error <- rnorm(n, mean = mean_error, sd = sigma_error)
      # } else {
      #   shape <- 4 / skew^2
      #   rate <- sqrt(shape) / sigma_error
      #   error <- rgamma(n, shape = shape, rate = rate)
      #   error <- error - shape / rate   # center to mean 0
      # }
if (abs(skew) < 0.01) {
  error <- rnorm(n, mean = mean_error, sd = sigma_error)
} else {
  # Convert desired mean=0, sd=3.4, and skewness into skew-normal parameters
  cp_params <- c(mean = mean_error, sd = sigma_error, skewness = skew)
  dp_params <- cp2dp(cp_params, family = "SN")
  
  error <- rsn(n, dp = dp_params)
}

# Optional: Quick check (you can remove later)
# cat("Actual mean:", round(mean(error), 4), " | Actual sd:", round(sd(error), 4), "\n")
      x <- runif(n, 0, 5)
      y <- beta0_original + beta1_original * x + error
      dat <- data.frame(x = x, y = y)
     
      ols <- lm(y ~ x, data = dat)
      slope[j] <- coef(ols)[2]
      
      se_slope[j] <- summary(ols)$coefficients["x", "Std. Error"]
      
      ci_lower <- slope[j] - 1.96 * se_slope[j]
      ci_upper <- slope[j] + 1.96 * se_slope[j]
      
      coverage[j] <- (ci_lower <= beta1_original) && (ci_upper >= beta1_original)
      
      bias[j]          <- (slope[j] - beta1_original) / beta1_original * 100
      interval_length[j] <- ci_upper - ci_lower
      power[j]         <- summary(ols)$coefficients["x", "Pr(>|t|)"] < 0.05
      
      absolute_percentage_bias[j] <- abs(slope[j] - beta1_original) / beta1_original * 100
    }
    
    results <- tibble(
    Run           = 1:repli,
    Sample_Size   = n,
    Model         = "OLS",
    Skewness = skew,
    Slope       = round(slope, 4),
    Bias = round(bias, 4),
    SE_Slope    = round(se_slope, 4),
    Coverage    = coverage,
    Power       = power,
    CI_Lower    = round(slope - 1.96 * se_slope),
    CI_Upper    = round(slope + 1.96 * se_slope),
    Interval_length = interval_length,
    True_Slope = round(beta1_original, 4)
  )
  ols_skew <- append(ols_skew, list(results))
  }
}

ols_skew <- bind_rows(ols_skew)

head(ols_skew, n = 3)
```

```{r eval = T, warning = F, message = F}
qr_skew <- list()

for (skew in skewness) {
  cat("\nSkewness =", skew, "\n")
  
  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
    
    slope <- numeric(repli)
    bias  <- numeric(repli)
    coverage <- logical(repli)
    se_slope         <- numeric(repli)
    interval_length <- numeric(repli)
    power <- logical(repli)
    absolute_percentage_bias <- numeric(repli)
    for (j in 1:repli) {
      # Error generation
      if (abs(skew) < 0.01) {
  error <- rnorm(n, mean = 0, sd = sigma_error)
} else {
  # Convert desired mean=0, sd=3.4, and skewness into skew-normal parameters
  cp_params <- c(mean = 0, sd = sigma_error, skewness = skew)
  dp_params <- cp2dp(cp_params, family = "SN")
  
  error <- rsn(n, dp = dp_params)
}
      
      x <- runif(n, 0, 5)
      y <- beta0_original + beta1_original * x + error
      dat <- data.frame(x = x, y = y)
      
      qr_fit <- rq(y ~ x, tau = 0.5, data = dat)
      
      slope[j] <- coef(qr_fit)[2]
      
      sum_qr <- summary(qr_fit, se = "boot", bsmethod = "mcmb")
      se_slope[j] <- sum_qr$coefficients["x", "Std. Error"]
      
      ci_lower <- slope[j] - 1.96 * se_slope[j]
      ci_upper <- slope[j] + 1.96 * se_slope[j]
      bias <- slope[j] - beta1_original
      coverage[j]     <- (ci_lower <= beta1_original) && (ci_upper >= beta1_original)
      interval_length[j] <- ci_upper - ci_lower
      power[j]        <- (ci_lower > 0) || (ci_upper < 0)   # 0 not in CI
      absolute_percentage_bias[j] <- abs(slope[j] - beta1_original) / beta1_original * 100
    }
    
    # Create results tibble
    results <- tibble(
      Run = 1:repli,
      Sample_Size = n,
      Model = "QR_tau_0.5",
      Skewness = skew,
      Slope = slope,
      Bias = round(bias, 4),
      SE_Slope = se_slope,
      Coverage = coverage,
      Power = power,
      CI_Lower = slope - 1.96 * se_slope,
      CI_Upper = slope + 1.96 * se_slope,
      Interval_length = interval_length,
      True_Slope = beta1_original
    )
    
    qr_skew <- append(qr_skew, list(results))
  }
}

# Combine OLS and QR
ols_qr_skew <- bind_rows(ols_skew, bind_rows(qr_skew))

# Quick view
ols_qr_skew[seq(1, nrow(ols_qr_skew), by = 1000), ]
```

```{r}
bias_skew <- ggplot(ols_qr_skew, aes(x = factor(Sample_Size), y = Bias, fill = Model)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  facet_wrap(~ Skewness, scales = "free_y", ncol = 3) +
  labs(title = "Bias of Slope Estimates", x = "Sample Size", fill = "Model") +
  theme_minimal()
print(bias_skew)
saveRDS(bias_skew, "bias_skew")
```

```{r}
slope_skew <- ggplot(ols_qr_skew, aes(x = factor(Sample_Size), y = Slope, fill = Model)) +
  geom_boxplot() +
  facet_wrap(~ Skewness, scales = "free_y") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
  labs(title = "Distribution of Slope Estimates")
print(bias_skew)
saveRDS(slope_skew, "slope_skew")
```

```{r}
power_coverage_skew <- ols_qr_skew %>%
  group_by(Skewness, Sample_Size, Model) %>%
  summarise(
    Power_rate    = mean(Power) * 100,
    Coverage_rate = mean(Coverage) * 100,
    .groups = "drop"
  )

power_skew <- ggplot(power_coverage_skew, aes(x = Sample_Size, y = Power_rate, color = Model)) +
  geom_line() +
  facet_wrap(~ Skewness, scales = "free_y") +
  geom_hline(yintercept = 95, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("rect", 
           xmin = -Inf, xmax = Inf,
           ymin = 95 - 1.96 * sqrt(95 * 0.95/repli),
           ymax = 95 + 1.96 * sqrt(95 * 0.95/repli),
           alpha = 0.1, fill = "blue") +
  labs(title = "Distribution of Power Rate of Skewed Data")

coverage_skew <- ggplot(power_coverage_skew, aes(x = Sample_Size, y = Coverage_rate, color = Model)) +
  geom_line() +
  facet_wrap(~ Skewness, scales = "free_y") +
  geom_hline(yintercept = 95, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("rect", 
           xmin = -Inf, xmax = Inf,
           ymin = 95 - 1.96 * sqrt(95 * 0.95/repli),
           ymax = 95 + 1.96 * sqrt(95 * 0.95/repli),
           alpha = 0.1, fill = "blue") +
  labs(title = "Distribution of Coverage Rate of Skewed Data")

saveRDS(power_skew, "power_skew")
saveRDS(coverage_skew, "coverage_skew")
power_skew
coverage_skew
```

## Type I error

```{r}
# set.seed(479)
# typeI_results <- list()
# for (skew in skewness) {
#   cat("\nSkewness =", skew, "\n")
#  
#   for (i in seq_along(sample_size)) {
#     n <- sample_size[i]
#    
#     reject_ols <- logical(repli)
#     reject_qr <- logical(repli)
#    
#     for (j in 1:repli) {
#       # Generate error
#       if (skew == 0) {
#         error <- rnorm(n, mean = 0, sd = 3.5)
#       } else {
#         shape <- 4 / skew^2
#         rate <- sqrt(shape) / 3.5
#         error <- rgamma(n, shape = shape, rate = rate)
#         error <- error - mean(error)
#       }
#      
#       x <- runif(n, 0, 5)
#       y <- beta0_original + 0 * x + error # TRUE slope = 0
#      
#       dat <- data.frame(x = x, y = y)
#      
#       # OLS
#       ols_fit <- lm(y ~ x, data = dat)
#       p_ols <- summary(ols_fit)$coefficients["x", "Pr(>|t|)"]
#       reject_ols[j] <- p_ols < 0.05
#      
#       # QR tau=0.5
#       qr_fit <- rq(y ~ x, tau = 0.5, data = dat)
#       sum_qr <- summary(qr_fit, se = "boot", bsmethod = "mcmb")
#       se_qr <- sum_qr$coefficients["x", "Std. Error"]
#       ci_l <- coef(qr_fit)[2] - 1.96 * se_qr
#       ci_u <- coef(qr_fit)[2] + 1.96 * se_qr
#       reject_qr[j] <- (ci_l > 0) || (ci_u < 0)
#     }
#    
#     # Calculate Type I error rate + 95% CI for OLS
#     p_ols <- mean(reject_ols)
#     se_ols <- sqrt(p_ols * (1 - p_ols) / repli)
#     ci_ols_lower <- p_ols - 1.96 * se_ols
#     ci_ols_upper <- p_ols + 1.96 * se_ols
#    
#     # Calculate Type I error rate + 95% CI for QR
#     p_qr <- mean(reject_qr)
#     se_qr <- sqrt(p_qr * (1 - p_qr) / repli)
#     ci_qr_lower <- p_qr - 1.96 * se_qr
#     ci_qr_upper <- p_qr + 1.96 * se_qr
#    
#     # Store results
#     typeI_results <- append(typeI_results, list(
#       tibble(
#         Sample_Size = n,
#         Skewness = skew,
#         Model = "OLS",
#         TypeI_Error = p_ols,
#         TypeI_Lower = ci_ols_lower,
#         TypeI_Upper = ci_ols_upper,
#         Rejections = sum(reject_ols)
#       )
#     ))
#    
#     typeI_results <- append(typeI_results, list(
#       tibble(
#         Sample_Size = n,
#         Skewness = skew,
#         Model = "QR_tau_0.5",
#         TypeI_Error = p_qr,
#         TypeI_Lower = ci_qr_lower,
#         TypeI_Upper = ci_qr_upper,
#         Rejections = sum(reject_qr)
#       )
#     ))
#   }
# }
# typeI_df <- bind_rows(typeI_results)
# # Display nice table
# typeI_df %>%
#   mutate(across(c(TypeI_Error, TypeI_Lower, TypeI_Upper), ~ round(.x * 100, 2))) %>%
#   select(Skewness, Sample_Size, Model, TypeI_Error, TypeI_Lower, TypeI_Upper, Rejections) %>%
#   arrange(Skewness, Sample_Size, Model)
```

```{r}
# set.seed(479)
# typeI_results <- list()
# 
# for (skew in skewness) {
#   cat("\nSkewness =", skew, "\n")
#  
#   for (i in seq_along(sample_size)) {
#     n <- sample_size[i]
#    
#     reject_ols <- logical(repli)
#     reject_qr  <- logical(repli)
#    
#     for (j in 1:repli) {
#       # Generate error
#       if (skew == 0) {
#         error <- rnorm(n, mean = 0, sd = 3.5)
#       } else {
#         shape <- 4 / skew^2
#         rate  <- sqrt(shape) / 3.5
#         error <- rgamma(n, shape = shape, rate = rate)
#         error <- error - mean(error)
#       }
#      
#       x <- runif(n, 0, 5)
#       y <- beta0_original + 0 * x + error # TRUE slope = 0
#      
#       dat <- data.frame(x = x, y = y)
#      
#       # OLS
#       ols_fit <- lm(y ~ x, data = dat)
#       p_ols <- summary(ols_fit)$coefficients["x", "Pr(>|t|)"]
#       reject_ols[j] <- p_ols < 0.05
#      
#       # QR tau=0.5
#       qr_fit <- rq(y ~ x, tau = 0.5, data = dat)
#       sum_qr <- summary(qr_fit, se = "boot", bsmethod = "mcmb")   # kept as requested
#       se_qr  <- sum_qr$coefficients["x", "Std. Error"]
#       ci_l   <- coef(qr_fit)[2] - 1.96 * se_qr
#       ci_u   <- coef(qr_fit)[2] + 1.96 * se_qr
#       reject_qr[j] <- (ci_l > 0) || (ci_u < 0)
#     }
#    
#     # Calculate Type I error rate + 95% CI 
#     p_ols <- mean(reject_ols)
#     se_ols <- sqrt(p_ols * (1 - p_ols) / repli)
#     ci_ols_lower <- p_ols - 1.96 * se_ols
#     ci_ols_upper <- p_ols + 1.96 * se_ols
#    
#     p_qr <- mean(reject_qr)
#     se_qr <- sqrt(p_qr * (1 - p_qr) / repli)
#     ci_qr_lower <- p_qr - 1.96 * se_qr
#     ci_qr_upper <- p_qr + 1.96 * se_qr
#    
#     # Store results - now side by side (one row per n & skew)
#     typeI_results <- append(typeI_results, list(
#       tibble(
#         Skewness      = skew,
#         Sample_Size   = n,
#         OLS_TypeI     = p_ols,
#         OLS_Lower     = ci_ols_lower,
#         OLS_Upper     = ci_ols_upper,
#         OLS_Rejections = sum(reject_ols),
#         QR_TypeI      = p_qr,
#         QR_Lower      = ci_qr_lower,
#         QR_Upper      = ci_qr_upper,
#         QR_Rejections = sum(reject_qr)
#       )
#     ))
#   }
# }
# 
# typeI_df <- bind_rows(typeI_results)
# 
# # Display nice table - OLS and QR side by side
# typeI_df %>%
#   select(Skewness, Sample_Size, OLS_TypeI, QR_TypeI) %>%
#   arrange(Skewness, Sample_Size)
#   print(typeI_df)
```

```{r}
set.seed(479)
typeI_results <- list()

for (skew in skewness) {
  cat("\nSkewness =", skew, "\n")
 
  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
   
    reject_ols <- logical(repli)
    reject_qr  <- logical(repli)
   
    for (j in 1:repli) {
      # Generate error
      if (skew == 0) {
        error <- rnorm(n, mean = 0, sd = 3.5)
      } else {
        shape <- 4 / skew^2
        rate  <- sqrt(shape) / 3.5
        error <- rgamma(n, shape = shape, rate = rate)
        error <- error - mean(error)
      }
     
      x <- runif(n, 0, 5)
      y <- beta0_original + 0 * x + error # TRUE slope = 0
     
      dat <- data.frame(x = x, y = y)
     
      # OLS
      ols_fit <- lm(y ~ x, data = dat)
      p_ols <- summary(ols_fit)$coefficients["x", "Pr(>|t|)"]
      reject_ols[j] <- p_ols < 0.05
     
      # QR tau=0.5
      qr_fit <- rq(y ~ x, tau = 0.5, data = dat)
      sum_qr <- summary(qr_fit, se = "boot", bsmethod = "mcmb")
      se_qr  <- sum_qr$coefficients["x", "Std. Error"]
      ci_l   <- coef(qr_fit)[2] - 1.96 * se_qr
      ci_u   <- coef(qr_fit)[2] + 1.96 * se_qr
      reject_qr[j] <- (ci_l > 0) || (ci_u < 0)
    }
   
    # Calculate Type I error rate + 95% CI for OLS
    p_ols <- mean(reject_ols)
    se_ols <- sqrt(p_ols * (1 - p_ols) / repli)
    ci_ols_lower <- p_ols - 1.96 * se_ols
    ci_ols_upper <- p_ols + 1.96 * se_ols
   
    # Calculate Type I error rate + 95% CI for QR
    p_qr <- mean(reject_qr)
    se_qr <- sqrt(p_qr * (1 - p_qr) / repli)
    ci_qr_lower <- p_qr - 1.96 * se_qr
    ci_qr_upper <- p_qr + 1.96 * se_qr
   
    # Store results - one row per model (as requested)
    typeI_results <- append(typeI_results, list(
      tibble(
        Skewness      = skew,
        Sample_Size   = n,
        Model         = "OLS",
        TypeI_Error   = p_ols,
        TypeI_Lower   = ci_ols_lower,
        TypeI_Upper   = ci_ols_upper,
        Rejections    = sum(reject_ols)
      )
    ))
    
    typeI_results <- append(typeI_results, list(
      tibble(
        Skewness      = skew,
        Sample_Size   = n,
        Model         = "QR_0.5",
        TypeI_Error   = p_qr,
        TypeI_Lower   = ci_qr_lower,
        TypeI_Upper   = ci_qr_upper,
        Rejections    = sum(reject_qr)
      )
    ))
  }
}

typeI_df <- bind_rows(typeI_results)

# Display nice table
typeI_df %>%
  select(Skewness, Sample_Size, Model, TypeI_Error, TypeI_Lower, TypeI_Upper, Rejections) %>%
  arrange(Skewness, Sample_Size, Model)
```

```{r}
# ggplot(typeI_df, aes(x = factor(Sample_Size), y = TypeI_Error, color = Model)) +
#   geom_point(size = 3) +
#   geom_errorbar(aes(ymin = TypeI_Lower, ymax = TypeI_Upper), width = 0.2) +
#   geom_hline(yintercept = 0.05, linetype = "dashed", color = "red", linewidth = 0.8) +
#   facet_wrap(~ Skewness, labeller = label_both) +
#   labs(title = "Type I Error Rate",
#        x = "Sample Size",
#        y = "Type I Error Rate") +
#   theme_minimal()
```

![](data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAACgAAAAaCAYAAADFTB7LAAAAcElEQVR4Xu3OwQmAQAxE0bClWYCW5N06tM6V2YPg5CjoF/JhLoHAi6iqn9eOefUbqrYvHY0cQDLyAlKRNyARmYA0ZMLRkAlGQyaU72tkAtlim7r/vJqDUDjlKBROOQyFU2icQuMUGqfQuBEaV1XPOwEx96nYACK8+wAAAABJRU5ErkJggg== "Run Current Chunk")

```{r}
ggplot(typeI_df, aes(x = factor(Sample_Size), y = TypeI_Error, color = Model)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = TypeI_Lower, ymax = TypeI_Upper), width = 0.25) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("rect", 
           xmin = -Inf, xmax = Inf,
           ymin = 0.05 - 1.96 * sqrt(0.05 * 0.95 / repli),
           ymax = 0.05 + 1.96 * sqrt(0.05 * 0.95 / repli),
           alpha = 0.1, fill = "blue") +
  facet_wrap(~ Skewness, labeller = label_both) +
  labs(title = "Type I Error Rate Simulation",
       x = "Sample Size",
       y = "Type I Error Rate") +
  theme_minimal()
```

```{r}

```
