---
title: "Data Analysis"
format:
  html:
    embed-resources: true 
    self-contained-math: true 
    html-math-method: mathjax
editor: visual
---

```{r eval = T, warning = F, message = F}
library(tidycensus)
library(tidyverse)
readRenviron("~/.Renviron")
library(tmap)
library(tigris)
library(sf)
library(spdep)
library(gstat)
library(nlme)
library(spatialreg)
library(ggplot2)
library(RColorBrewer)
library(patchwork)
library(GWmodel)
tmap_mode("view")
library(quantreg)
library(gamboostLSS)
library(caret)
library(e1071)
options(tigris_use_cache = TRUE)
```

[https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/county/time-series/CA-037/tavg/8/12/1895-2024?base_prd=true&begbaseyear=1901&endbaseyear=2000&utm](https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/county/time-series/CA-037/tavg/8/12/1895-2024?base_prd=true&begbaseyear=1901&endbaseyear=2000&utm_source=chatgpt.com)

<https://aqs.epa.gov/aqsweb/airdata/download_files.html#Annual>

[Public Use Microdata Survey Variable Explanation](https://www2.census.gov/programs-surveys/acs/tech_docs/pums/data_dict/PUMS_Data_Dictionary_2024.pdf)

```{r eval = T, warning = F, message = F}
# aqi <- read.csv("~/Downloads/annual_aqi_by_county_2023.csv")
# geo_fl <- get_acs(geography = "county", year = 2023, 
#                   output = "wide", geometry = TRUE,
#                   variables = c(
#                     median_income = "B19013_001E",
#                     poverty_total = "B17001_001E",
#                     poverty_bad = "B17001_002E",
#                     bachelor = "B15003_022E",
#                     master = "B15003_023",
#                     doctor = "B15003_025",
#                     professional = "B15003_024",
#                     education = "B15003_001E",
#                     unemployed = "B23025_005E",
#                     labor_force = "B23025_003E",
#                     not_in_force = "B23025_007E"
#                   ))
```

JWMNP Numeric 3 Travel time to work bbb .N/A (not a worker or worker who worked from home) 1..200 .1 to 200 minutes to get to work (Top-coded)

JWTRNS Character 2 Means of transportation to work bb .N/A (not a worker-not in the labor force, including persons .under 16 years; unemployed; employed, with a job but not at .work; Armed Forces, with a job but not at work) 01 .Car, truck, or van 02 .Bus 03 .Subway or elevated rail 04 .Long-distance train or commuter rail 05 .Light rail, streetcar, or trolley 06 .Ferryboat 07 .Taxi or ride-hailing services 08 .Motorcycle 09 .Bicycle 10 .Walked 11 .Worked from home 12 .Other method

COW Character 1 Class of worker b .N/A (less than 16 years old/NILF who last worked more than 5 .years ago or never worked) 1 .Employee of a private for-profit company or business, or of an .individual, for wages, salary, or commissions 2 .Employee of a private not-for-profit, tax-exempt, or .charitable organization 3 .Local government employee (city, county, etc.) 4 .State government employee 5 .Federal government employee 6 .Self-employed in own not incorporated business, professional .practice, or farm 7 .Self-employed in own incorporated business, professional .practice or farm 8 .Working without pay in family business or farm 9 .Unemployed and last worked 5 years ago or earlier or never .worked

SSIP Numeric 5 Supplementary Security Income past 12 months (use ADJINC to adjust SSIP to constant dollars) bbbbb .N/A (less than 15 years old) 0 .None 4..30000 .\$4 to \$30000 (Rounded and top-coded) SSP Numeric 5 Social Security income past 12 months (use ADJINC to adjust SSP to constant dollars) bbbbb .N/A (less than 15 years old) 0 .None 4..55000 .\$4 to \$55000 (Rounded and top-coded) WAGP Numeric 6 Wages or salary income past 12 months (use ADJINC to adjust WAGP to constant dollars) bbbbbb .N/A (less than 15 years old) 0 .None 4..999999 .\$4 to 999999 (Rounded and top-coded)

WAGP Numeric 6 Wages or salary income past 12 months (use ADJINC to adjust WAGP to constant dollars) bbbbbb .N/A (less than 15 years old) 0 .None 4..999999 .\$4 to 999999 (Rounded and top-coded)

Model:

```{r eval = T, warning = F, message = F}
pums_data <- get_pums(variables = c(
  "PUMA",
  "JWMNP",     # Commute time
  "JWTRNS",    # Mean of travel
  "SCHL",      # Educational attainment
  "MAR",       # Marital status
  "PAOC",      # Presence and age of own children (proxy for young <6)
  "WAGP",      # Wage and salary income
  "WKHP",      # Usual hours worked per week
  "WKWN",      # Weeks worked in past 12 months
  "ADJINC"
  ),
  state = "FL",           
  survey = "acs5",        
  year = 2023, 
  recode = TRUE
)
```

```{r eval = T, warning = F, message = F}
clean_data <- pums_data %>%
  filter(
    !is.na(JWMNP),
    JWTRNS == "03",
    WAGP > 0,
    WKHP >= 1,
    WKWN >= 1
  ) %>% 
  mutate(
    ADJINC = as.numeric(as.character(ADJINC)),
    adj_wagp = WAGP * (ADJINC / 1000000),
    hourly_wage = adj_wagp / (WKWN * WKHP),
    ln_wage = log(hourly_wage)
  ) %>%
  mutate(
    has_child = factor(case_when(
      PAOC %in% c(
        "With own children under 6 years only",
        "With own children under 6 years and 6 to 17 years",
        "With own children 6 to 17 years only"
      ) ~ "Has Child",
      PAOC == "No own children under 18 years" ~ "No Child",
      TRUE ~ NA_character_
    )),
    has_young_child = factor(case_when(
      PAOC %in% c(
        "With own children under 6 years only",
        "With own children under 6 years and 6 to 17 years"
      ) ~ "Has Young Child (<6)",
      PAOC %in% c(
        "No own children under 6 years",
        "With own children 6 to 17 years only"
      ) ~ "No Young Child",
      TRUE ~ NA_character_
    ))
  ) %>%
  mutate(
    educ_level = case_when(
      SCHL %in% c(
        "No schooling completed", "Nursery school, preschool",
        "Kindergarten", "Grade 1", "Grade 2", "Grade 3", "Grade 4",
        "Grade 5", "Grade 6", "Grade 7", "Grade 8", "Grade 9",
        "Grade 10", "Grade 11", "12th grade - no diploma"
      ) ~ "Less than HS",
      SCHL %in% c(
        "Regular high school diploma", "GED or alternative credential",
        "Some college, but less than 1 year", "1 or more years of college credit, no degree",
        "Associate's degree"
      ) ~ "HS or Some College",
      SCHL == "Bachelor's degree" ~ "Bachelor's",
      SCHL %in% c(
        "Master's degree", "Professional degree beyond a bachelor's degree",
        "Doctorate degree"
      ) ~ "Graduate",
      TRUE ~ "Other"
    ) %>% factor(),
    married = factor(ifelse(
      MAR %in% c("Married", "Married, spouse absent"),
      "Married",
      "Not Married"
    ))
  ) 
# %>%
#   filter(
#     hourly_wage > quantile(hourly_wage, 0.1, na.rm = TRUE),
#     hourly_wage < quantile(hourly_wage, 0.9, na.rm = TRUE)
#   )
```

```{r}
ggplot(clean_data, aes(x = JWMNP)) +
  geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "steelblue", color = "white") +
  geom_density(color = "red", linewidth = 1.5) +
  labs(
    title = "Distribution of Commute Time",
    x = "Commute Time (minutes)",
    y = "Density"
  ) +
  theme_minimal()

ggplot(clean_data, aes(x = WAGP)) +
  geom_histogram(bins = 60, fill = "steelblue", color = "black") +
  labs(title = "Income Distribution", x = "Wage and Salary Income", y = "Count") +
  theme_minimal()
```

```{r eval = T, warning = F, message = F}
commute_by_puma <- clean_data %>%
  group_by(PUMA) %>%
  summarise(mean_commute = mean(JWMNP, na.rm = TRUE), .groups = "drop")

acs_puma <- get_acs(
  geography = "public use microdata area",
  variables = c(medincome = "B19013_001"),
  state = "FL",
  year = 2023,
  survey = "acs5"
)

merged <- clean_data %>%
  left_join(acs_puma, by = c("PUMA" = "GEOID"))
```

```{r eval = T, warning = F, message = F}
ggplot(merged) + geom_sf(aes(fill = JWMNP)) +
  scale_fill_gradientn(colors = brewer.pal(9, 'Purples')) +
  theme_minimal()
```

```{r}
commute_by_puma <- clean_data %>%
  group_by(PUMA) %>%
  summarise(
    mean_commute = mean(JWMNP, na.rm = TRUE),
    median_commute = median(JWMNP, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

fl_pumas <- pumas(
  state = "FL",
  year = 2023,
  class = "sf"
)

map_data <- fl_pumas %>%
  mutate(PUMA = as.character(PUMACE20)) %>%
  left_join(commute_by_puma, by = "PUMA")

ggplot(map_data) +
  geom_sf(aes(fill = mean_commute), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(na.value = "grey90") +
  labs(
    title = "Average Commute Time by PUMA in Florida",
    fill = "Minutes"
  ) +
  theme_minimal()

ggplot(map_data) +
  geom_sf(aes(fill = median_commute), color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(na.value = "grey90") +
  labs(
    title = "Median Commute Time by PUMA in Florida",
    fill = "Minutes"
  ) +
  theme_minimal()
```

```{r eval = T, warning = F, message = F}
# data_name = read.csv(url("https://hbiostat.org/data/repo/bodyfat.csv"))

pm25_fl <- read_csv("~/Downloads/ad_viz_plotval_data.csv")
ozone_fl <- read_csv("~/Downloads/ad_viz_plotval_data (1).csv")
length(unique(ozone_fl$County))
```

```{r eval = T, warning = F, message = F}
plot(data$`Daily Mean PM2.5 Concentration`,data$Date)
```

```{r eval = T, warning = F, message = F}


```
