---
title: "Untitled"
format: html
editor: visual
---

# Simulate data for Linear quantile regression

<https://stats.stackexchange.com/questions/499149/monte-carlo-simulation-for-quantile-regression#:~:text=n%20=%201000%20beta0%20=%202.1;,Add%20a%20comment%20%7C>

[How to simulate data](https://stirlingcodingclub.github.io/simulating_data/index.html#:~:text=Literature%20Cited-,Introduction:%20Simulating%20data,statistical%20analyses%20and%20data%20visualisation.)

[Design simulation scenarios for quantile](https://scholarcommons.sc.edu/cgi/viewcontent.cgi?params=/context/etd/article/5508/&path_info=Howard_sc_0202M_15604.pdf)

[Simulation of quantile regression of heteroscedastic error variance](https://www.academia.edu/90825129/The_Simulation_Study_to_Test_the_Performance_of_Quantile_Regression_Method_With_Heteroscedastic_Error_Variance#:~:text=6)%20Equivalently%2C%20we%20may%20rewrite,small%20bias%20and%20small%20variance.)

Yanuar, F. (2017). The Simulation Study to Test the Performance of Quantile Regression Method With Heteroscedastic Error Variance. CAUCHY: Jurnal Matematika Murni Dan Aplikasi. <https://doi.org/10.18860/CA.V5I1.4209>

In this section, I simulated a data set to compare the linear regression with quantile regression. I started off by creating random predictors values and picked a linear function: $y_0 = 4 + 0.2 \times x$. In this simulation, the heteroscedasticity is created by letting the errors have a positive correlation to $x$: $\epsilon = 0.3 + 0.8\times x$, thus the dispersion of y increases as x increases.

<https://stirlingcodingclub.github.io/simulating_data/index.html#:~:text=Literature%20Cited-,Introduction:%20Simulating%20data,statistical%20analyses%20and%20data%20visualisation.>

<https://stats.stackexchange.com/questions/499149/monte-carlo-simulation-for-quantile-regression#:~:text=n%20=%201000%20beta0%20=%202.1;,Add%20a%20comment%20%7C>

## Heteroskedastic error with increasing variance simulation

```{r eval = T, warning = F, message = F}
# Library section
library(quantreg)
library(ggplot2)
```

```{r eval = T, warning = F, message = F}
# Some functions
# MSE
# rmse_func <- function(true, pred) {sqrt((mean(pred - true))^2)}
sample_size <- c(20, 50, 100, 200, 300, 500, 650)
slope_func <- function(data, indices) {
  d <- data[indices, ]
  fit <- lm(y ~ x, data = d)
  coef(fit)[2]
}
```

```{r eval = T, warning = F, message = F}
set.seed(479)
# List of needed metrics
coverage <- numeric(length(sample_size)) #set coverage with the same obs as sample size
n_sizes <- length(sample_size)
mean_slope <- numeric(n_sizes)
sd_slope <- numeric(n_sizes)
mean_se_slope  <- numeric(n_sizes)
power <- numeric(n_sizes)

for (i in seq_along(sample_size)) {
  chosen_size <- sample_size[i]

  coverage.estimate <- logical(1000)
  slope <- numeric(1000)
  se_slope <- numeric(1000)
  error <- numeric(1000)
  beta1.true <- 0.4
  power_count <- logical(1000)

  for (rep in 1:1000) {
    x <- runif(chosen_size, min = 1, max = 6)
    y <- 15 + 0.4 * x + rnorm(chosen_size, 0, 2.3) * (1 + x/14)

    # OLS
    ols <- lm(y ~ x)
    confidence_interval <- confint(ols, "x", level = 0.95)
    coverage.estimate[rep] <- (confidence_interval[1] <= beta1.true) && (confidence_interval[2] >= beta1.true)
    slope[rep] <- coef(ols)[2]
    se_slope[rep] <- summary(ols)$coefficients["x", "Std. Error"]
    power_count[rep] <- (summary(ols)$coefficients["x", "Pr(>|t|)"] < 0.05)
  }

  coverage[i] <- mean(coverage.estimate) * 100
  mean_slope[i] <- mean(slope)
  mean_se_slope[i]     <- mean(se_slope)
  sd_slope[i] <- sd(slope)
  power[i] <- mean(power_count)*100
}
results <- data.frame(
  n              = sample_size,
  Mean_Slope     = round(mean_slope, 4),
  SE_Slope       = round(mean_se_slope, 4),
  Coverage_pct   = round(coverage, 1),
  POWER = round(power, 1)
)

print("OLS")
print(results)
```

```{r eval = T, warning = F, message = F}
set.seed(479)

beta_original <- 0.4
eps_sig <- 2.3
b <- 1/14
repli <- 1000
# Quantile regression
taus <- c(.25, .5, .75, .9)
n_sizes <- length(sample_size)
dat <- data.frame()

results <- list()

for (tau in taus) {
  cat("Quantile =", tau,"th\n")
  beta1.true <- beta_original + qnorm(tau) * eps_sig * b
  cat("True slope =", beta1.true,"\n")
  
  
  coverage <- numeric(n_sizes)
  power <- numeric(n_sizes)
  mean_slope <- numeric(n_sizes)
  sd_slope <- numeric(n_sizes)
  mean_se_slope  <- numeric(n_sizes)
  mae <- numeric(n_sizes)

  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
    coverage.estimate <- logical(repli)
    slope <- numeric(repli)
    se_slope <- numeric(repli)
    error <- numeric(repli)
    power_count <- logical(repli)

    for (rep in 1:repli) {
      x <- runif(n, min = 1, max = 6)
    
      y <- 15 + 0.4 * x + rnorm(n, 0, 2.3) * (1 + x/14)
      # y <- 34 + beta1.true * x + rnorm(chosen_size, 0, 2.9) *(1 + 0.9*x)
      dat <- data.frame(x = c(x), y = c(y))

      # QR
      qr <- rq(y~ x, tau = tau)
   
      s <- summary(qr, se = "boot", B = 50)
      slope[rep] <- coef(qr)[2]
      se_slope[rep] <- s$coefficients["x", "Std. Error"]
      
      ci_lower <- slope[rep] - 1.96 * se_slope[rep]
      ci_upper <- slope[rep] + 1.96 * se_slope[rep]

      coverage.estimate[rep] <- (ci_lower <= beta1.true) && (ci_upper >= beta1.true)
      error[rep] <- mean(abs(dat$y - predict(qr, newdata = dat)))
      power_count[rep] <- !(ci_lower <= 0 && ci_upper >= 0)
      }
    
    coverage[i] <- mean(coverage.estimate) * 100
    mean_slope[i] <- mean(slope)
    mean_se_slope[i] <- mean(se_slope)
    sd_slope[i] <- sd(slope)
    mae[i] <- (mean(error))
    power[i] <- mean(power_count) * 100
  }
  
  results_tau <- data.frame(n = sample_size,
                            Mean_Slope  = round(mean_slope, 4),
                            SE_Slope  = round(mean_se_slope, 4),
                            MAE_QR  = round(mae, 4),
                            Coverage_pct  = round(coverage, 1),
                            POWER = round(power, 1)
                            )
  print(results_tau)
  cat("\n")
}
```

## Skewed with constant variance simulation

```{r eval = T, warning = F, message = F}
# library(sn) # library of skew to normal dsitribution
set.seed(479)
skewness <- c(0, 0.5, 1, 2, 3, 6)
repli <- 500

beta0_original <- 15
beta1_original <- 0.4

# y <- beta0_original + beta1_original * x + error

sample_length <- length(sample_size)


for (skew in skewness) {
  cat("\nSkewness =", skew, "\n")
  
  avg_slope <- numeric(sample_length)
  avg_bias_percent <- numeric(sample_length)
  coverage_percent <- numeric(sample_length)
  avg_interval_length <- numeric(sample_length)
  power_percent <- numeric(sample_length)
  mae <- numeric(sample_length)

  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
    slope <- numeric(repli)
    bias <- numeric(repli)
    coverage.estimate <- logical(repli)
    interval_length <- numeric(repli)
    power <- logical(repli)
    absolute_bias <- numeric(repli)
    se_slope <- numeric(repli)
    
    for (j in 1:repli) {
      # Error formula 
      if (skew == 0) {
        error <- rnorm(n, 0, 2.5)
      } else if (skew == 0.5) {
        error <- rgamma(n, shape = 16, rate = 1.6)
        error <- error - 16/1.6
      } else if (skew == 1) {
        error <- rgamma(n, shape = 4, rate = 0.8)
        error <- error - 4/0.8
      } else if (skew == 2) {
        error <- rgamma(n, shape = 1, rate = 0.4)
        error <- error - 1/0.4
      } else if (skew == 3) {
        error <- rgamma(n, shape = 0.45, rate = 0.27)
        error <- error - 0.45/0.27
      } else if (skew == 6) {
        error <- rgamma(n, shape = 0.12, rate = 0.14)
        error <- error - 0.12/0.14
      } else {
        stop("Invalid ratio")
      }
      
      x <- runif(n, 0, 2)
      y <- beta0_original + beta1_original * x + error
      dat <- data.frame(x = x, y = y)
      
      ols <- lm(y ~ x, data = dat)
      slope[j] <- coef(ols)[2]
      
      bias[j] <- (slope[j] - beta1_original) / beta1_original * 100
      
      bootstrap <- boot::boot(data = dat, statistic = slope_func, R = 50)
      confidence_interval <- quantile(bootstrap$t, probs = c(0.025, 0.975))
      coverage.estimate[j] <- (confidence_interval[1] <= beta1_original) && (confidence_interval[2] >= beta1_original)
      
      interval_length[j] <- confidence_interval[2] - confidence_interval[1]
      power[j] <- !(confidence_interval[1] <= 0 && confidence_interval[2] >= 0)
      
      absolute_bias[j] <- abs(slope[j] - beta1_original) / beta1_original * 100
      se_slope[j] <- sd(bootstrap$t)
    
    }
    avg_slope[i] <- mean(slope)
    avg_bias_percent[i] <- mean(bias)
    coverage_percent[i] <- mean(coverage.estimate) * 100
    avg_interval_length[i] <- mean(interval_length)
    power_percent[i] <- mean(power) * 100
    mae[i] <- mean(absolute_bias)
  }

  results <- data.frame(n = sample_size,
                        avg_slope = round(avg_slope, 4),
                        bias_percent = round(avg_bias_percent, 4),
                        avg_interval_leng = round(avg_interval_length, 4),
                        coverage_per = round(coverage_percent, 4),
                        power = round(power_percent, 2),
                        mae = round(mae, 4)
                        )
  print("OLS")
  print(results)
}
```

```{r eval = T, warning = F, message = F}
set.seed(479)

for (skew in skewness) {
  cat("\nSkewness =", skew, "\n")

  avg_slope <- numeric(sample_length)
  avg_bias_percent <- numeric(sample_length)
  coverage_percent <- numeric(sample_length)
  avg_interval_length <- numeric(sample_length)
  power_percent <- numeric(sample_length)
  mae <- numeric(sample_length)

  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
    slope <- numeric(repli)
    bias <- numeric(repli)
    coverage <- logical(repli)
    interval_length <- numeric(repli)
    power <- logical(repli)
    absolute_bias <- numeric(repli)

    for (j in 1:repli) {
      if (skew == 0) {
        error <- rnorm(n, 0, 2.5)
      } else if (skew == 0.5) {
        error <- rgamma(n, shape = 16, rate = 1.6)
        error <- error - 16/1.6
      } else if (skew == 1) {
        error <- rgamma(n, shape = 4, rate = 0.8)
        error <- error - 4/0.8
      } else if (skew == 2) {
        error <- rgamma(n, shape = 1, rate = 0.4)
        error <- error - 1/0.4
      } else if (skew == 3) {
        error <- rgamma(n, shape = 0.45, rate = 0.27)
        error <- error - 0.45/0.27
      } else if (skew == 6) {
        error <- rgamma(n, shape = 0.12, rate = 0.14)
        error <- error - 0.12/0.14
      } else {
        stop("Invalid ratio")
      }

      x <- runif(n, 0, 2)
      y <- beta0_original + beta1_original * x + error
      dat <- data.frame(x = x, y = y)

      qr_fit <- rq(y ~ x, tau = 0.5, data = dat)
      s <- summary(qr_fit, se = "boot", B = 50)

      slope[j] <- coef(qr_fit)[2]
      bias[j] <- (slope[j] - beta1_original) / beta1_original * 100

      se_x <- s$coefficients["x", "Std. Error"]
      ci_lower <- slope[j] - 1.96 * se_x
      ci_upper <- slope[j] + 1.96 * se_x

      coverage[j] <- (ci_lower <= beta1_original) && (ci_upper >= beta1_original)
      interval_length[j] <- ci_upper - ci_lower
      power[j] <- s$coefficients["x", "Pr(>|t|)"] < 0.05
      absolute_bias[j] <- abs(slope[j] - beta1_original) / beta1_original * 100
    }

    avg_slope[i] <- mean(slope)
    avg_bias_percent[i] <- mean(bias)
    coverage_percent[i] <- mean(coverage) * 100
    avg_interval_length[i] <- mean(interval_length)
    power_percent[i] <- mean(power) * 100
    mae[i] <- mean(absolute_bias)
  }

  results <- data.frame(
    n = sample_size,
    avg_slope = round(avg_slope, 4),
    bias_percent = round(avg_bias_percent, 4),
    avg_interval_leng = round(avg_interval_length, 4),
    coverage_per = round(coverage_percent, 4),
    power = round(power_percent, 2),
    mae = round(mae, 4)
  )

  print("QR")
  print(results)
}
```

```{r}
set.seed(479)

# Type I Error Simulation for QR (Median)
cat("=== Type I Error Rate for Median Quantile Regression ===\n")
sample_size <- c(20, 200, 650)
typeI_results <- list()

for (skew in skewness) {
  cat("\nSkewness =", skew, "\n")
  
  typeI_rate <- numeric(length(sample_size))
  rejections <- numeric(length(sample_size))
  
  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
    reject <- logical(repli)
    
    for (j in 1:repli) {
      # Generate centered errors (mean ≈ 0)
      if (skew == 0) {
        error <- rnorm(n, mean = 0, sd = 2.5)
      } else if (skew == 0.5) {
        error <- rgamma(n, shape = 16, rate = 1.6)
      } else if (skew == 1) {
        error <- rgamma(n, shape = 4, rate = 0.8)
      } else if (skew == 2) {
        error <- rgamma(n, shape = 1, rate = 0.4)
      } else if (skew == 3) {
        error <- rgamma(n, shape = 0.45, rate = 0.27)
      } else if (skew == 6) {
        error <- rgamma(n, shape = 0.12, rate = 0.14)
      } else {
        stop("Invalid skewness value")
      }
      
      # Center the gamma errors (better than hard-coded mean)
      if (skew != 0) error <- error - mean(error)
      
      x <- runif(n, 0, 2)
      y <- 0 + 0 * x + error          # True slope = 0 (intercept can be 0)
      
      dat <- data.frame(x = x, y = y)
      
      # Median Quantile Regression
      qr_fit <- rq(y ~ x, tau = 0.5, data = dat)
      
      s <- summary(qr_fit, se = "boot", bsmethod = "xy", R = 100)
      
      coef_table <- s$coefficients
      est_slope  <- coef_table["x", "Value"]
      se_slope   <- coef_table["x", "Std. Error"]
      
      # Normal approximation CI
      ci_lower <- est_slope - 1.96 * se_slope
      ci_upper <- est_slope + 1.96 * se_slope
      
      reject[j] <- (ci_lower > 0) || (ci_upper < 0)
    }
    
    typeI_rate[i] <- mean(reject)
    rejections[i] <- sum(reject)
  }
  
  temp_df <- data.frame(
    Skewness     = skew,
    n            = sample_size,
    TypeI_Error  = round(typeI_rate, 4),
    TypeI_Pct    = round(typeI_rate * 100, 2),
    Rejections   = rejections,
    stringsAsFactors = FALSE
  )
  
  typeI_results[[as.character(skew)]] <- temp_df
  print(temp_df)
}

# Combine results
final_typeI <- do.call(rbind, typeI_results)
rownames(final_typeI) <- NULL

cat("\n=== Final Type I Error Summary ===\n")
print(final_typeI)
```

```{r}
library(quantreg)   # for rq()
library(stats)      # lm() is base, but good to be explicit

set.seed(479)

# Assuming these are defined in previous chunks:
# skewness <- c(0, 0.5, 1, 2, 3, 6)
# repli     <- 1000   # or higher

cat("=== Type I Error Simulation: Median QR vs OLS ===\n")

sample_size <- c(20, 200, 650)
typeI_results <- list()

for (skew in skewness) {
  cat("\nSkewness =", skew, "\n")
  
  typeI_qr  <- numeric(length(sample_size))
  typeI_ols <- numeric(length(sample_size))
  rejections_qr  <- numeric(length(sample_size))
  rejections_ols <- numeric(length(sample_size))
  
  for (i in seq_along(sample_size)) {
    n <- sample_size[i]
    reject_qr  <- logical(repli)
    reject_ols <- logical(repli)
    
    for (j in 1:repli) {
      # Generate centered errors
      if (skew == 0) {
        error <- rnorm(n, mean = 0, sd = 2.5)
      } else if (skew == 0.5) {
        error <- rgamma(n, shape = 16, rate = 1.6)
      } else if (skew == 1) {
        error <- rgamma(n, shape = 4, rate = 0.8)
      } else if (skew == 2) {
        error <- rgamma(n, shape = 1, rate = 0.4)
      } else if (skew == 3) {
        error <- rgamma(n, shape = 0.45, rate = 0.27)
      } else if (skew == 6) {
        error <- rgamma(n, shape = 0.12, rate = 0.14)
      } else {
        stop("Invalid skewness value")
      }
      
      if (skew != 0) error <- error - mean(error)
      
      x <- runif(n, 0, 2)
      y <- 15 + 0 * x + error
      
      dat <- data.frame(x = x, y = y)
      
      # ====================== Median Quantile Regression ======================
      qr_fit <- rq(y ~ x, tau = 0.5, data = dat)
      
      tryCatch({
        s_qr <- summary(qr_fit, se = "boot", bsmethod = "xy", R = 1000)
        
        coef_qr   <- s_qr$coefficients
        est_qr    <- coef_qr["x", "Value"]
        se_qr     <- coef_qr["x", "Std. Error"]
        
        ci_lower_qr <- est_qr - 1.96 * se_qr
        ci_upper_qr <- est_qr + 1.96 * se_qr
        
        reject_qr[j] <- (ci_lower_qr > 0) || (ci_upper_qr < 0)
        
      }, error = function(e) {
        reject_qr[j] <- FALSE   # conservative: treat failure as non-rejection
      })
      
      # ====================== Ordinary Least Squares (OLS) ======================
      ols_fit <- lm(y ~ x, data = dat)
      s_ols   <- summary(ols_fit)
      
      coef_ols   <- s_ols$coefficients
      est_ols    <- coef_ols["x", "Estimate"]
      se_ols     <- coef_ols["x", "Std. Error"]
      
      # Use t-critical value (more accurate than 1.96 for small n)
      t_crit     <- qt(0.975, df = n - 2)
      ci_lower_ols <- est_ols - t_crit * se_ols
      ci_upper_ols <- est_ols + t_crit * se_ols
      
      reject_ols[j] <- (ci_lower_ols > 0) || (ci_upper_ols < 0)
    }
    
    # Store rates
    typeI_qr[i]  <- mean(reject_qr)
    typeI_ols[i] <- mean(reject_ols)
    rejections_qr[i]  <- sum(reject_qr)
    rejections_ols[i] <- sum(reject_ols)
  }
  
  # Create table for this skewness
  temp_df <- data.frame(
    Skewness      = skew,
    n             = sample_size,
    OLS_TypeI     = round(typeI_ols, 4),
    QR_TypeI      = round(typeI_qr, 4),
    # QR_Pct        = round(typeI_qr * 100, 2),
    # OLS_Pct       = round(typeI_ols * 100, 2),
    # Rej_QR        = rejections_qr,
    # Rej_OLS       = rejections_ols,
    stringsAsFactors = FALSE
  )
  
  typeI_results[[as.character(skew)]] <- temp_df
  print(temp_df)
}

# Final combined results
final_typeI <- do.call(rbind, typeI_results)
rownames(final_typeI) <- NULL

cat("\n=== Final Type I Error Summary (QR vs OLS) ===\n")
print(final_typeI)
```
