---
title: "Quantile Regression in Practice: Simulation and an Application to Stroke Prevalence"
subtitle: "IDC6940 Capstone"
author: "Ngoc Bui (Advisor: Dr. Cohen)"
date: '`r Sys.Date()`'
format:
  revealjs
course: Capstone Projects in Data Science
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

## Introduction {style="font-size: 0.6em"}

- This is a comparison study between Original Least Squares (OLS) vs. Quantile Regression (QR) [@koenker2005].
- Though usually we want to know the average outcome of the target variable, we are more interested in the behavior of the lower 40% or upper 20% of it. 
- For example, researchers would be more interested in the lower percentile of children's stunting distribution to assess the malnutrition situation [@Ashortstoryonhowandwhy].
- Regression models:
  - OLS: $OLS = \alpha + \beta \times X_i^T + \epsilon$
  - QR: $Q(\tau) = \alpha_{\tau} + \beta_{\tau} \times X_i^T + \epsilon_{\tau}$
- In QR, different percentiles have different intercepts, slopes, and errors.
- Loss functions:
  - OLS: $\epsilon = min(\sum (y_i - \beta x_i^T)^2)$
  - QR: $\epsilon_\tau = min(\tau \times \sum_{y_i > Q(\tau)} |y_i - Q(\tau)| + (1 - \tau) \times \sum_{y_1 < Q(\tau)} |y_i - Q(\tau) |)$

# Simulation
## Simulation {.smaller}
- The simulation is a replication of [@simulation].
    - First case: Skewed residual distribution.
    - Second case: Residual Heteroskedastic Variance.
- For every predictor x is uniformly generated from 0 to 5, its response y is $y_i = 20 + 0.6\times x_i + \epsilon_i \sim N(0, 1.9)$
- The non-normal behaviors are added in the residual distribution, which directly create non-normal data.
- This study employs a Monte Carlo simulation [@Montecarlomath2019] to simulate 1000 replications/modeling counts of OLS vs. QR on a linear model.
- Sample size = 50, 100, 200, 450, 700.


## Comparing Metrics {style="font-size: 0.6em"}
- The power percentage is defined by the proportion of replications where the null hypothesis, the slope of the regression is non-significant, is rejected [@muthen2002use].
$$
\begin{aligned}
Power &= \sum_{1}^{replications} I(p\_value < 0.05), \text{ for } I(p\_value) \\
&=
\begin{cases}
1 & \text{if } p\_value < 0.05 \\
0 & \text{otherwise}
\end{cases}
\end{aligned}
$$


- Coverage percentage is the proportion of the replications that the true slope falls inside the 95% confidence interval [@muthen2002use].
$$
Coverage = \sum_{1}^{replications} I(LB_{95} < slope < UP_{95})
$$

- Type I Error Rate is defined exactly as the Power Percentage but based on the true slope is 0 [@muthen2002use].
$$
\text{Type I error} = \sum_{1}^{replications} I(p\_value < 0.05)
$$


## Skewed residual distribution {.smaller}
- Main comparison: OLS vs. QR at median.
- Residuals have different levels of skewness while maintaining the $\mu$ at 0 and $\sigma$ at 1.9.
- Skewness levels: 0, 2.5, 3, 4, 4.5, 5.
- To creat skewness in the residuals, Gamma distribution is applied since it has simple code in R while adding the skewness.
$$
f(x_i) = \frac{x_i^{\alpha - 1} e^{-x_i/\beta}}{\Gamma(\alpha)\,\beta^{\alpha}}, \quad \text{where } \Gamma(\alpha) = \int_0^\infty x^{\alpha -1} e^{-x} \, dx
$$
- To generate the skewness in R, the relationship between the skewness, standard deviation, shape and rate parameters are:
$$
\begin{aligned}
Skewness &= \frac{2}{\sqrt(shape)} = \frac{2}{\sqrt(\alpha)} \\
STD &= \sqrt(shape) \times rate = \sqrt(\alpha) \times \beta \\
\end{aligned}
$$

## Skewed residual distribution

```{r}
#| fig-cap: "Distribution of Power Percentage between OLS and QR (r = 0.5) at different skewness levels"
#| label: fig-power_skew
power_skew <- readRDS("power_skew.rds")
power_skew
```


## Skewed residual distribution {.smaller}
```{r}
#| fig-cap: "Coverage Percentage between OLS and QR(r = 0.5) at different skewness levels"
#| label: fig-cover_skew
cover_skew <- readRDS("coverage_skew.rds")
cover_skew
```


## Skewed residual distribution {.smaller}
```{r}
#| fig-cap: "Type I Error between OLS and QR(r = 0.5) at different skewness levels"
#| label: fig-type1-qr
type1_qr <- readRDS("typei_2models_skew.rds")
type1_qr
```


## Residual Heteroskedastic Variance {.smaller}
- The heteroskedastic variance is generated by multiplying the original residuals by a linear equation: 
$\epsilon(x_i) = N(0, 1.9) \times (1 + x_i \times \frac{1}{7})$.
- Percentile level: $\tau^{th} = 25, 50, 75, 90$.

```{r eval = T, warning = F, message = F}
#| fig-cap: "The change in slopes at different quantile between homoscedastic and heteroscedastic response for n = 700"
#| label: fig-homo-heter1

library(patchwork)
homo_var1 <- readRDS("homo_var.rds")
heter_var1 <- readRDS("heter_var.rds")
p1 <- homo_var1 | heter_var1
p1
```


## Residual Heteroskedastic Variance {.smaller}
- Percentiles: $\tau^{th} = 25, 50, 75, 90$.
```{r}
#| fig-cap: "Power and Coverage Percentage between OLS and QRs on heteroscedastic response"
#| label: fig-power-cover-heter
power_heter <- readRDS("power_heter.rds")
cover_heter <- readRDS("cover_heter.rds")
power_heter | cover_heter
```


## Residual Heteroskedastic Variance {.smaller}
```{r}
#| fig-cap: "Type I Error between OLS and QRs on heteroscedastic response"
#| label: fig-typei-heter
typei_2models_he <- readRDS("typei_2models_he.rds")
typei_2models_he
```

# Application

## Stroke Prevalence in the US (2022) {.smaller}
**Application**: County-level stroke prevalence in the United States (2022)
- With the increase number of stroke cases, identifying the factors affecting stroke prevalence help inform health practitioner and the community with planning and applying healthy practice.

- The factors come from social practice or living environment. The data is from CDC and County Health Rankings.

- The chosen predictors are:
  - smoke: Adult smoking prevalence — percentage of adults who are current smokers.

  - diabetes_prev: Diabetes prevalence — percentage of adults aged 18 and older with diagnosed diabetes.
  
  - obesity_prev: Adult obesity prevalence — percentage of adults with a body mass index ≥ 30 kg/m².
  
  - air_pollution: Average daily density of fine particulate matter (PM2.5) in µg/m³.
  
  - physical_inactive: Physical inactivity prevalence — percentage of adults who reported no leisure-time physical activity.

## 
```{r}
#| fig-cap: "Histogram of Stroke Prevalence (left) and Stroke Prevalence by Diabete Prevalence on OLS vs QRs (right)"
#| label: fig-plot.dist
plot.dist <- readRDS("plot.dist.rds")
hist.stroke <- readRDS("hist.stroke.rds")
p2 <- hist.stroke | plot.dist
p2
```

## Results {.smaller}
- OLS: STROKE ~ 0.58352 + 10.68295(smoke) + 11.50501(diabetes_prev) + 0.85461(physical_inactive)
- QR(median): STROKE ~ 0.47272 + 10.78209(smoke) + 9.00408(diabetes_prev) + 1.82025(physical_inactive)
```{r}
#| fig-cap: "Slope Estimates between OLS vs. QR on Different Quantile Levels"
#| label: fig-coef.stroke
coef.stroke <- readRDS("coef.stroke.rds")
coef.stroke
```

## Comparing Prediction Intervals
- Prediction interval is explained in [@heskes1996practical].
- We want the n% of real observations to be equal to the size of the interval.
- For 70% prediction interval of OLS and between predicted $15^{th}$ and $85^{th}$ percentile of the QR models, we want approximately 70% of the observed data fall within this range.
- Prediction intervals: 50%, 70%, 80%, 90%.


## Comparison of Prediction Intervals {.smaller}
- As the prediction interval increases, the OLS shows signs of over coverage, since 99.26% of the real stroke data fall in the 90% prediction interval.
- QRs show closer fit of the number of real observations and the prediction intervals regardless.

```{r}
#| tbl-cap: "Prediction Interval between OLS and QRs on Stroke Prevalence"
#| label: tab-typei-heter
pre.interval.stroket <- readRDS("pre.interval.stroket.rds")
pre.interval.stroket
```


## Conclusion
- OLS produced one value of slope estimates regardless of the skewness and different effects of the predictors in different percentiles.
- QR displays better coverage in different levels of prediction intervals.
- As the prediction intervals increase, OLS overly covers the real number of observations actually in the intervals.
- There are other predictors that can affect stroke outcomes that are not included in this study.

\begin{center}
Thank You!
\end{center}

```{r}

```

## References


